### 2016-11-28 reports/Effects_of_AWD_N_on_SB.rmd
Initial commit
Add data;

Add README.md files;

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/Development/svm-r-markdown-templates/svm-latex-ms.tex
title: "Effects of Alternate Wetting and Drying Irrigation and Nitrogen Fertilization on Sheath Blight of Rice"
thanks: "Replication files are available on the author's Github account (http://github.com/adamhsparks/AWD_N_SB). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: Adam.Sparks@usq.edu.au"
author:
- name: Adam H. Sparks
  affiliation: Centre for Crop Health, University of Southern Queensland, West St., Toowoomba, QLD 4350, Australia
- name: Nancy P. Castilla
  affiliation: Crop and Environmental Sciences Division, International Rice Research Institute, Los Baños, Laguna 4031, Philippines
abstract: "Abstract limited to 200 words in one paragraph."
keywords: "*Rhizoctonia solani*"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt

# spacing: double
bibliography: ~/Dropbox/master.bib
biblio-style: apsr
---

Introduction goes here with no heading

# Materials and Methods

# Results

## Data analysis

\pagebreak

# Discussion

# Acknowledgements

<!--
# Literature Cited
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\noindent
-->