### 2017-04-14 Analysis_nested_effects.Rmd
Add nested random effects analysis
Analyse the nested efffect of the main plot with replicate

Also name R chunks in both Rmd files

---
title: "Analysis with nested effects"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```
## Notes on this analysis before starting

These analyses nest the main plot effect as a random effect with the replicate.

#### How to interpret these analyses

For a quick overview, see this thread on Stack Overflow that walks through the output of `summary.MCMCglmm`,
which is used below. [R: Making sense of the output of a MCMCglmm](http://stackoverflow.com/questions/20993643/r-making-sense-of-the-output-of-a-mcmcglmm)

The MCMCglmm documentation is more in-depth but should still be referenced.
```{r MCMCglmm, eval=FALSE}
library(MCMCglmm)
vignette("Overview")
vignette("CourseNotes")
```

#### Changes between 2015 and 2016

Before the analysis, note that due to changes between the years, the analysis
must be carried out on each year separately. The 2015 data and 2016 data cannot
be combined due to changes in inoculation methods; there are other changes too,
but the main one is this. Therefore, the analyses will be conducted separately for
leaf severity and Leaf severity for 2015 dry season and 2016 dry season.
Comparisons will only be observational and cannot be statistically compared.

#### Iterations
Fitting the models with the default iterations resulted in some less than
acceptable models. Because of this, the number of iterations has been increased,
`nitt = 5e+05` and `burnin = 5000` and `thin = 100`. This results in much
smaller errors for random effects and better models.

## 2015
### 2015 Leaf Sheath Blight Severity Model
```{r 2015_LShB}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_lmm_2015 <- MCMCglmm(LShB_AUDPS ~ WMGT * NRTE,
                          random = ~REP:WMGT, 
                          data = as.data.frame(AUDPS_2015),
                          verbose = FALSE,
                          prior = eprior,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          pr = TRUE)

summary(LShB_lmm_2015)

# create data frames for generating diagnostic plots
reps <- data.frame(LShB_lmm_2015$Sol[, c(1, 7:14)])
x <- 1:nrow(reps)
reps <- melt(reps)

trts <-  data.frame(LShB_lmm_2015$Sol[, 2:6])
names(trts)[names(trts) == "WMGTFLD.NRTE100"] <- "FLD.NRTE100"
names(trts)[names(trts) == "WMGTFLD.NRTE120"] <- "FLD.NRTE120"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(LShB_lmm_2015$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2015 Diagnostic Plots for Replicates, Leaf ShB Severity")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2015 Reps Replicate Posteriors for Replicates, Leaf ShB Severity")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2015 Diagnostic Plots for Treatments, Leaf ShB Severity")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2015 Posteriors for Treatments, Leaf ShB Severity")

# check random effects
plotTrace(LShB_lmm_2015$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(LShB_lmm_2015$VCV)
plot_joint_random_error_dist_nested(d = rdf,
                             title = ("2015 Random Error Distribution for Leaf ShB Severity"))
```

### 2015 Tiller Sheath Blight Severity Model
```{r 2015_TShB}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_severity_lmm_2015 <- MCMCglmm(TShB_AUDPS ~ WMGT * NRTE,
                                   random = ~REP:WMGT, 
                          data = as.data.frame(AUDPS_2015),
                          verbose = FALSE,
                          prior = eprior,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          pr = TRUE)

summary(TShB_severity_lmm_2015)

# create data frames for generating diagnostic plots
reps <- data.frame(TShB_severity_lmm_2015$Sol[, c(1, 7:14)])
reps <- melt(reps)

trts <-  data.frame(TShB_severity_lmm_2015$Sol[, 2:6])
names(trts)[names(trts) == "WMGTFLD.NRTE100"] <- "FLD.NRTE100"
names(trts)[names(trts) == "WMGTFLD.NRTE120"] <- "FLD.NRTE120"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(TShB_severity_lmm_2015$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2015 Diagnostic Plots for Replicates, Tiller ShB Severity")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2015 Reps Replicate Posteriors for Replicates, Tiller ShB Severity")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2015 Diagnostic Plots for Treatments, Tiller ShB Severity")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2015 Posteriors for Treatments, Tiller ShB Severity")

# check random effects
plotTrace(TShB_severity_lmm_2015$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(TShB_severity_lmm_2015$VCV)
plot_joint_random_error_dist_nested(d = rdf,
                             title = ("2015 Random Error Distribution for Tiller ShB Severity"))
```

### 2015 Tiller Sheath Blight Incidence Model
```{r 2015_TShB_incidence}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_incidence_lmm_2015 <- MCMCglmm(TShB_incidence_mean ~ WMGT * NRTE,
                                    random = ~REP:WMGT, 
                          data = as.data.frame(AUDPS_2015),
                          verbose = FALSE,
                          prior = eprior,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          pr = TRUE)

summary(TShB_incidence_lmm_2015)

# create data frames for generating diagnostic plots
reps <- data.frame(TShB_incidence_lmm_2015$Sol[, c(1, 7:14)])
reps <- melt(reps)

trts <-  data.frame(TShB_incidence_lmm_2015$Sol[, 2:6])
names(trts)[names(trts) == "WMGTFLD.NRTE100"] <- "FLD.NRTE100"
names(trts)[names(trts) == "WMGTFLD.NRTE120"] <- "FLD.NRTE120"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(TShB_incidence_lmm_2015$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2015 Diagnostic Plots for Replicates, Tiller ShB Incidence")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2015 Reps Replicate Posteriors for Replicates, Tiller ShB Incidence")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2015 Diagnostic Plots for Treatments, Tiller ShB Incidence")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2015 Posteriors for Treatments, Tiller ShB Incidence")

# check random effects
plotTrace(TShB_incidence_lmm_2015$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(TShB_incidence_lmm_2015$VCV)
plot_joint_random_error_dist_nested(d = rdf,
                             title = ("2015 Random Error Distribution for Tiller ShB Incidence"))
```

*******

## 2016

### 2016 Leaf Sheath Blight Severity Model
```{r 2016_LShB}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_lmm_2016 <- MCMCglmm(LShB_AUDPS ~ WMGT * NRTE,
                          random = ~REP:WMGT, 
                          data = as.data.frame(AUDPS_2016),
                          verbose = FALSE,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          prior = eprior,
                          pr = TRUE)

summary(LShB_lmm_2016)

# create data frames for generating diagnostic plots
reps <- data.frame(LShB_lmm_2016$Sol[, c(1, 5:12)])
reps <- melt(reps)

trts <-  data.frame(LShB_lmm_2016$Sol[, 2:4])
names(trts)[names(trts) == "WMGTFLD.NRTE180"] <- "FLD.NRTE180"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(LShB_lmm_2016$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2016 Diagnostic Plots for Replicates, Leaf ShB Severity")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2016 Reps Replicate Posteriors for Replicates, Leaf ShB Severity")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2016 Diagnostic Plots for Treatments, Leaf ShB Severity")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2016 Posteriors for Treatments, Leaf ShB Severity")

# check random effects
plotTrace(LShB_lmm_2016$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(LShB_lmm_2016$VCV)
plot_joint_random_error_dist_nested(d = rdf,
                             title = ("2016 Random Error Distribution for Leaf ShB Severity"))
```

### 2016 Tiller Sheath Blight Severity Model
```{r 2016_TShB}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_severity_lmm_2016 <- MCMCglmm(TShB_AUDPS ~ WMGT * NRTE,
                                   random = ~REP:WMGT, 
                          data = as.data.frame(AUDPS_2016),
                          verbose = FALSE,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          prior = eprior,
                          pr = TRUE)

summary(TShB_severity_lmm_2016)

# create data frames for generating diagnostic plots
reps <- data.frame(TShB_severity_lmm_2016$Sol[, c(1, 5:12)])
reps <- melt(reps)

trts <-  data.frame(TShB_severity_lmm_2016$Sol[, 2:4])
names(trts)[names(trts) == "WMGTFLD.NRTE180"] <- "FLD.NRTE180"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(TShB_severity_lmm_2016$Sol)
# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2016 Diagnostic Plots for Replicates, Tiller ShB Severity")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2016 Reps Replicate Posteriors for Replicates, Tiller ShB Severity")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2016 Diagnostic Plots for Treatments, Tiller ShB Severity")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2016 Posteriors for Treatments, Tiller ShB Severity")

# check random effects
plotTrace(TShB_severity_lmm_2016$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(TShB_severity_lmm_2016$VCV)
plot_joint_random_error_dist_nested(d = rdf,
                             title = ("2016 Random Error Distribution for Tiller ShB Severity"))
```

### 2016 Tiller Sheath Blight Incidence Model
```{r 2016_TShB_incidence}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_incidence_lmm_2016 <- MCMCglmm(TShB_incidence_mean ~ WMGT * NRTE, 
                                    random = ~REP:WMGT, 
                          data = as.data.frame(AUDPS_2016),
                          verbose = FALSE,
                          prior = eprior,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          pr = TRUE)

summary(TShB_incidence_lmm_2016)

# create data frames for generating diagnostic plots
reps <- data.frame(TShB_incidence_lmm_2016$Sol[, c(1, 5:12)])
reps <- melt(reps)

trts <-  data.frame(TShB_incidence_lmm_2016$Sol[, 2:4])
names(trts)[names(trts) == "WMGTFLD.NRTE100"] <- "FLD.NRTE100"
names(trts)[names(trts) == "WMGTFLD.NRTE120"] <- "FLD.NRTE120"
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(TShB_incidence_lmm_2016$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2016 Diagnostic Plots for Replicates, Tiller ShB Incidence")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2016 Reps Replicate Posteriors for Replicates, Tiller ShB Incidence")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2016 Diagnostic Plots for Treatments, Tiller ShB Incidence")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2016 Posteriors for Treatments, Tiller ShB Incidence")

# check random effects
plotTrace(TShB_incidence_lmm_2016$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(TShB_incidence_lmm_2016$VCV)
plot_joint_random_error_dist_nested(d = rdf,
                             title = ("2016 Random Error Distribution for Tiller ShB Incidence"))
```

# Conclusions

The models all appear to be good fits after increasing the number of iterations.
None of the treatments in either year, 2015 or 2016, were significant. The
see `pMCMC` values and also posterior graphs. There is a large amount of overlap
in all of the posterior graphs.