### 2017-04-26 Probability_distribution_checks.Rmd
Finalise data cleaning and run analysis on AUDPS only
Finalised data cleaning (again). Data are now in a tidy data frame for
ordinal analysis. AUDPS has been generated from the tiller incidence
data and analysed using MCMCglmm

All raw data graphs have been updated

---
title: "Probability Distribution Checks"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```

Here we check the probability distribution that best fits these data. I'm following a method from http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html. As you see in that guide, there are several distributions that can be checked, however, the data both fit a normal distribution.

## 2015 Check probability distribution fit for tiller sheath blight incidence

```{r, results='hide', message=FALSE, warning=FALSE}
AUDPS15 <- subset(AUDPS, YEAR == "2015")
f1 <- fitdist(AUDPS15$AUDPS, "norm")
plot(f1)
car::qqp(AUDPS15$AUDPS, "norm")

```

*******

## 2016 Check probability distribution fit for tiller sheath blight incidence

```{r, results='hide', message=FALSE, warning=FALSE}
AUDPS156 <- subset(AUDPS, YEAR == "2016")
f1 <- fitdist(AUDPS156$AUDPS, "norm")
plot(f1)
car::qqp(AUDPS156$AUDPS, "norm")

```

## Conclusions

To read the graphs, we want to select a distribution where most of the points fall between the dotted red lines. On the y-axis are the observations and on the x-axis the quantiles modelled by the distribution. A perfect distribution fit is represented by the solid red line and the dashed red lines are the confidence intervals of the perfect distribution fit. Here both sets of data fit a normal distribution.