### 2017-05-10 Probability_distribution_checks.Rmd
Update probability distribution checks

---
title: "Probability Distribution Checks"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```

Here we check the probability distribution that best fits these data. I'm following a method from http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html. As you see in that guide, there are several distributions that can be checked, however, the data both fit a normal distribution.

## 2015 Check probability distribution fit for tiller sheath blight incidence

### Normal
```{r norm, results='hide', message=FALSE, warning=FALSE}
AUDPS15 <- subset(AUDPS, YEAR == "2015")
f1 <- fitdist(AUDPS15$AUDPS, "norm")
plot(f1)
car::qqp(AUDPS15$AUDPS, "norm")
```

### Exponential

```{r exp, results='hide', message=FALSE, warning=FALSE}
AUDPS15 <- subset(AUDPS, YEAR == "2015")
f1 <- fitdist(AUDPS15$AUDPS, "exp")
plot(f1)
car::qqp(AUDPS15$AUDPS, "exp")
```

*******

## 2016 Check probability distribution fit for tiller sheath blight incidence

Until the issue with more diseased tillers than tillers in a hill is addressed,
this section will not work completely and properly.

```{r, results='hide', message=FALSE, warning=FALSE}
AUDPS156 <- subset(AUDPS, YEAR == "2016")
# f1 <- fitdist(AUDPS156$AUDPS, "norm")
#plot(f1)
car::qqp(AUDPS156$AUDPS, "norm")

```

## Conclusions

To read the graphs, we want to select a distribution where most of the points fall between the dotted red lines. On the y-axis are the observations and on the x-axis the quantiles modelled by the distribution. A perfect distribution fit is represented by the solid red line and the dashed red lines are the confidence intervals of the perfect distribution fit. Here both sets of data fit a normal distribution.