### 2017-08-01 test_analysis.Rmd
Add test analysis with lme4; clean up Analysis.Rmd file

---
  title: "Analysis of split plot design using lme4"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```

## Notes on This Analysis Before Starting

The base levels for this analysis are:

- 2015 `NRTE:0`

- 2015 `WMGT:FLD`

- 2016 `NRTE:60`

- 2016 `WMGT:FLD`

#### Changes between 2015 and 2016

Before the analysis, note that due to changes between the years, the analysis
must be carried out on each year separately. The 2015 data and 2016 data cannot
be combined due to changes in inoculation methods; there are other changes too,
but the main one is this. Therefore, the analyses will be conducted separately
such that comparisons will only be observational and cannot be statistically
compared.

## Setup

The `AUDPS` object is created when the project is loaded, in the file [munge/03_preprocess_data.R](./munge/03_preprocess_data.R). However, because it is a `tibble` and the treatments exist in a single column for graphing the raw data, this object needs a few minor changes to be usable for the analysis.

First, separate the TRT column into the two treatments for analysis.

```{r data_setup}
AUDPS <- separate(data = AUDPS, col = TRT, sep = "_", into = c("WMGT", "NRTE"))
AUDPS <- mutate_at(.tbl = AUDPS, .funs = factor, .vars = c("WMGT", "NRTE"))
```

Now create individual data frames for the analysis.

```{r create_analysis_df}
AUDPS_2015 <- as.data.frame(AUDPS[AUDPS$YEAR == 2015, ])
AUDPS_2015 <- droplevels(AUDPS_2015)

# relevel factors for easier interpretation of  analysis
AUDPS_2015 <- within(AUDPS_2015, NRTE <- relevel(NRTE, ref = "N0"))
AUDPS_2015 <- within(AUDPS_2015, WMGT <- relevel(WMGT, ref = "FLD"))

AUDPS_2016 <- as.data.frame(AUDPS[AUDPS$YEAR == 2016, ])
AUDPS_2016 <- droplevels(AUDPS_2016)

# relevel factors for easier interpretation of  analysis
AUDPS_2016 <- within(AUDPS_2016, NRTE <- relevel(NRTE, ref = "N60"))
AUDPS_2016 <- within(AUDPS_2016, WMGT <- relevel(WMGT, ref = "FLD"))
```

Now that the `AUDPS_2015` and `AUDPS_2016` `data.frames` exist, we can start the analysis.

## 2015

### 2015 Tiller Sheath Blight Incidence Model

```{r 2015_TShB_incidence}
library(pbkrtest)
lmodr <- lmer(TShB_percent_AUDPS ~ WMGT * NRTE + (1|REP), data = AUDPS_2015)
lmoda <- lmer(TShB_percent_AUDPS ~ WMGT + NRTE + (1|REP), data = AUDPS_2015)

KRmodcomp(lmodr, lmoda)

lmodi <- lmer(yield ~ irrigation + (1|field),data=irrigation)
KRmodcomp(lmoda, lmodi)

lmodv <- lmer(yield ~  variety + (1|field),data=irrigation)
KRmodcomp(lmoda, lmodv)
```

### 2015 Tiller Sheath Blight Severity Model

```{r 2015_TShB_severity}
eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_severity_lmm_2015 <- MCMCglmm(TShB_percent_AUDPS ~ WMGT * NRTE,
                                   random = ~REP,
                                   data = AUDPS_2015,
                                   verbose = FALSE,
                                   prior = eprior,
                                   nitt = 5e+05,
                                   burnin = 5000,
                                   thin = 100,
                                   pr = TRUE)

summary(TShB_severity_lmm_2015)

# create data frames for generating diagnostic plots
reps <- data.frame(TShB_severity_lmm_2015$Sol[, c(1, 7:10)])
reps <- melt(reps)

trts <-  data.frame(TShB_severity_lmm_2015$Sol[, 2:6])
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(TShB_severity_lmm_2015$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2015 Diagnostic Plots for Replicates, Tiller ShB Severity")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2015 Reps Replicate Posteriors for Replicates, Tiller ShB Severity")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2015 Diagnostic Plots for Treatments, Tiller ShB Severity")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2015 Posteriors for Treatments, Tiller ShB Severity")

# check random effects
plotTrace(TShB_severity_lmm_2015$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(TShB_severity_lmm_2015$VCV)
plot_joint_random_error_dist(d = rdf,
                             title = ("2015 Random Error Distribution for Tiller ShB Severity"))
```

### 2015 Leaf Sheath Blight Severity Model

```{r 2015_LShB_severity}
eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_severity_lmm_2015 <- MCMCglmm(LShB_percent_AUDPS ~ WMGT * NRTE,
                                   random = ~REP,
                                   data = AUDPS_2015,
                                   verbose = FALSE,
                                   prior = eprior,
                                   nitt = 5e+05,
                                   burnin = 5000,
                                   thin = 100,
                                   pr = TRUE)

summary(LShB_severity_lmm_2015)

# create data frames for generating diagnostic plots
reps <- data.frame(LShB_severity_lmm_2015$Sol[, c(1, 7:10)])
reps <- melt(reps)

trts <-  data.frame(LShB_severity_lmm_2015$Sol[, 2:6])
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(LShB_severity_lmm_2015$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2015 Diagnostic Plots for Replicates, Leaf ShB Severity")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2015 Reps Replicate Posteriors for Replicates, Leaf ShB Severity")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2015 Diagnostic Plots for Treatments, Leaf ShB Severity")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2015 Posteriors for Treatments, Leaf ShB Severity")

# check random effects
plotTrace(LShB_severity_lmm_2015$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(LShB_severity_lmm_2015$VCV)
plot_joint_random_error_dist(d = rdf,
                             title = ("2015 Random Error Distribution for Leaf ShB Severity"))
```

*******

  ## 2016

  ### 2016 Tiller Sheath Blight Incidence Model

  ```{r 2016_TShB_incidence}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))

TShB_incidence_lmm_2016 <- MCMCglmm(TShB_inc_AUDPS ~ WMGT * NRTE,
                                    random = ~REP,
                                    data = AUDPS_2016,
                                    verbose = FALSE,
                                    prior = eprior,
                                    nitt = 5e+05,
                                    burnin = 5000,
                                    thin = 100,
                                    pr = TRUE)

summary(TShB_incidence_lmm_2016)

# create data frames for generating diagnostic plots
reps <- data.frame(TShB_incidence_lmm_2016$Sol[, c(1, 5:8)])
reps <- melt(reps)

trts <-  data.frame(TShB_incidence_lmm_2016$Sol[, 2:4])
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(TShB_incidence_lmm_2016$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2016 Diagnostic Plots for Replicates, Tiller ShB Incidence")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2016 Reps Replicate Posteriors for Replicates, Tiller ShB Incidence")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2016 Diagnostic Plots for Treatments, Tiller ShB Incidence")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2016 Posteriors for Treatments, Tiller ShB Incidence")

# check random effects
plotTrace(TShB_incidence_lmm_2016$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(TShB_incidence_lmm_2016$VCV)
plot_joint_random_error_dist(d = rdf,
                             title = ("2016 Random Error Distribution for Tiller ShB Incidence"))
```

### 2016 Tiller Sheath Blight Severity Model

```{r 2016_TShB_severity}
eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_severity_lmm_2016 <- MCMCglmm(TShB_percent_AUDPS ~ WMGT * NRTE,
                                   random = ~REP,
                                   data = AUDPS_2016,
                                   verbose = FALSE,
                                   prior = eprior,
                                   nitt = 5e+05,
                                   burnin = 5000,
                                   thin = 100,
                                   pr = TRUE)

summary(TShB_severity_lmm_2016)

# create data frames for generating diagnostic plots
reps <- data.frame(TShB_severity_lmm_2016$Sol[, c(1, 5:8)])
reps <- melt(reps)

trts <-  data.frame(TShB_severity_lmm_2016$Sol[, 2:4])
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(TShB_severity_lmm_2016$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2016 Diagnostic Plots for Replicates, Tiller ShB Severity")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2016 Reps Replicate Posteriors for Replicates, Tiller ShB Severity")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2016 Diagnostic Plots for Treatments, Tiller ShB Severity")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2016 Posteriors for Treatments, Tiller ShB Severity")

# check random effects
plotTrace(TShB_severity_lmm_2016$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(TShB_severity_lmm_2016$VCV)
plot_joint_random_error_dist(d = rdf,
                             title = ("2016 Random Error Distribution for Tiller ShB Severity"))
```

### 2016 Leaf Sheath Blight Severity Model

```{r 2016_LShB_severity}
eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_severity_lmm_2016 <- MCMCglmm(LShB_percent_AUDPS ~ WMGT * NRTE,
                                   random = ~REP,
                                   data = AUDPS_2016,
                                   verbose = FALSE,
                                   prior = eprior,
                                   nitt = 5e+05,
                                   burnin = 5000,
                                   thin = 100,
                                   pr = TRUE)

summary(LShB_severity_lmm_2016)

# create data frames for generating diagnostic plots
reps <- data.frame(LShB_severity_lmm_2016$Sol[, c(1, 5:8)])
reps <- melt(reps)

trts <-  data.frame(LShB_severity_lmm_2016$Sol[, 2:4])
trts <- melt(trts)

# Create a dummy x-axis variable for plotting
x <- 1:nrow(LShB_severity_lmm_2016$Sol)

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2016 Diagnostic Plots for Replicates, Leaf ShB Severity")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2016 Reps Replicate Posteriors for Replicates, Leaf ShB Severity")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2016 Diagnostic Plots for Treatments, Leaf ShB Severity")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2016 Posteriors for Treatments, Leaf ShB Severity")

# check random effects
plotTrace(LShB_severity_lmm_2016$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(LShB_severity_lmm_2016$VCV)
plot_joint_random_error_dist(d = rdf,
                             title = ("2016 Random Error Distribution for Leaf ShB Severity"))
```

### Model Fit

The models all appear to be good fits.

None of the diagnostic plots show any signs of auto-correlation or any other obvious patterns.

The random effects all appear to be acceptable, the dotted line stays near to the solid line with no discernible patterns.

## Conclusions

### Tiller Sheath Blight Incidence

In 2015 the highest N rate, N120, was significantly different from the base, N0 rate.

In 2016 both the AWD and highest N rate, N180, were significantly different from the FLD and N60 treatments.

No interactions were found to be significant in the tiller sheath blight incidence.

## Tiller Sheath Blight Severity

In 2015 both the N100 and N120 rates were significantly different from the N0 treatment. The interaction of the N100 rate with AWD was also significantly different than N0 and FLD treatments.

In 2016 the N180 was significantly different from the N60 rate and the interaction of the N180 and AWD was significantly different than the N60 and FLD interaction.

### Leaf Sheath Blight Severity

In 2015 the N120 was significantly different than the N0 treatment. The interactions of both the N100 and N120 with AWD were significantly different than the N0 FLD interaction.

In 2016 the N180 was significantly different from the N60 treatment. The AWD was significantly different from the FLD and the interaction of N180:AWD was significantly different from the N60:FLD treatments interaction.

## R Session Info

```{r session_info, echo=FALSE}
devtools::session_info()
```