### 2017-08-03 lme4_analysis.Rmd
Clean up summary output of lmer

---
title: "Analysis of split plot design using lme4"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```

## Notes on This Analysis Before Starting

The base levels for this analysis are:

- 2015 `NRTE:0`

- 2015 `WMGT:FLD`

- 2016 `NRTE:60`

- 2016 `WMGT:FLD`

#### Changes between 2015 and 2016

Before the analysis, note that due to changes between the years, the analysis
must be carried out on each year separately. The 2015 data and 2016 data cannot
be combined due to changes in inoculation methods; there are other changes too,
but the main one is this. Therefore, the analyses will be conducted separately
such that comparisons will only be observational and cannot be statistically
compared.

## Setup

The `AUDPS` object is created when the project is loaded, in the file [munge/03_preprocess_data.R](./munge/03_preprocess_data.R). However, because it is a `tibble` and the treatments exist in a single column for graphing the raw data, this object needs a few minor changes to be usable for the analysis.

First, separate the TRT column into the two treatments for analysis.

```{r data_setup}
AUDPS <- separate(data = AUDPS, col = TRT, sep = "_", into = c("WMGT", "NRTE"))
AUDPS <- mutate_at(.tbl = AUDPS, .funs = factor, .vars = c("WMGT", "NRTE"))
```

Now create individual data frames for the analysis.

```{r create_analysis_df}
AUDPS_2015 <- as.data.frame(AUDPS[AUDPS$YEAR == 2015, ])
AUDPS_2015 <- droplevels(AUDPS_2015)

# relevel factors for easier interpretation of  analysis
AUDPS_2015 <- within(AUDPS_2015, NRTE <- relevel(NRTE, ref = "N0"))
AUDPS_2015 <- within(AUDPS_2015, WMGT <- relevel(WMGT, ref = "FLD"))

AUDPS_2016 <- as.data.frame(AUDPS[AUDPS$YEAR == 2016, ])
AUDPS_2016 <- droplevels(AUDPS_2016)

# relevel factors for easier interpretation of  analysis
AUDPS_2016 <- within(AUDPS_2016, NRTE <- relevel(NRTE, ref = "N60"))
AUDPS_2016 <- within(AUDPS_2016, WMGT <- relevel(WMGT, ref = "FLD"))
```

Now that the `AUDPS_2015` and `AUDPS_2016` `data.frames` exist, we can start the analysis.

## 2015

### 2015 Tiller Sheath Blight Incidence Model

```{r 2015_TShB_incidence}
library(lme4)
library(lsmeans)
TShB_inc_AUDPS <- lmer(TShB_inc_AUDPS ~ WMGT * NRTE + (1|REP), data = AUDPS_2015)
print(summary(TShB_inc_AUDPS), correlation = FALSE)

plot(TShB_inc_AUDPS)
qqnorm(resid(TShB_inc_AUDPS))
qqline(resid(TShB_inc_AUDPS))

print(lsmeans(TShB_inc_AUDPS, list(pairwise ~ NRTE)), adjust = c("tukey"))
print(lsmeans(TShB_inc_AUDPS, list(pairwise ~ WMGT)), adjust = c("tukey"))
print(lsmeans(TShB_inc_AUDPS, list(pairwise ~ WMGT | NRTE)), adjust = c("tukey"))
```

### 2015 Tiller Sheath Blight Severity Model

```{r 2015_TShB_severity}
TShB_sev_AUDPS <- lmer(TShB_percent_AUDPS ~ WMGT * NRTE + (1|REP), data = AUDPS_2015)
print(summary(TShB_sev_AUDPS), correlation = FALSE)

plot(TShB_sev_AUDPS)
qqnorm(resid(TShB_sev_AUDPS))
qqline(resid(TShB_sev_AUDPS))

print(lsmeans(TShB_sev_AUDPS, list(pairwise ~ NRTE)), adjust = c("tukey"))
print(lsmeans(TShB_sev_AUDPS, list(pairwise ~ WMGT)), adjust = c("tukey"))
print(lsmeans(TShB_sev_AUDPS, list(pairwise ~ WMGT | NRTE)), adjust = c("tukey"))
```

### 2015 Leaf Sheath Blight Severity Model

```{r 2015_LShB_severity}
LShB_sev_AUDPS <- lmer(LShB_percent_AUDPS ~ WMGT * NRTE + (1|REP), data = AUDPS_2015)
print(summary(LShB_sev_AUDPS), correlation = FALSE)

plot(LShB_sev_AUDPS)
qqnorm(resid(LShB_sev_AUDPS))
qqline(resid(LShB_sev_AUDPS))

print(lsmeans(LShB_sev_AUDPS, list(pairwise ~ NRTE)), adjust = c("tukey"))
print(lsmeans(LShB_sev_AUDPS, list(pairwise ~ WMGT)), adjust = c("tukey"))
print(lsmeans(LShB_sev_AUDPS, list(pairwise ~ WMGT | NRTE)), adjust = c("tukey"))
```

*******

## 2016

### 2016 Tiller Sheath Blight Incidence Model

```{r 2016_TShB_incidence}
TShB_inc_AUDPS <- lmer(TShB_inc_AUDPS ~ WMGT * NRTE + (1|REP), data = AUDPS_2016)
print(summary(TShB_inc_AUDPS), correlation = FALSE)

plot(TShB_inc_AUDPS)
qqnorm(resid(TShB_inc_AUDPS))
qqline(resid(TShB_inc_AUDPS))

print(lsmeans(TShB_inc_AUDPS, list(pairwise ~ NRTE)), adjust = c("tukey"))
print(lsmeans(TShB_inc_AUDPS, list(pairwise ~ WMGT)), adjust = c("tukey"))
print(lsmeans(TShB_inc_AUDPS, list(pairwise ~ WMGT | NRTE)), adjust = c("tukey"))
```

### 2016 Tiller Sheath Blight Severity Model

```{r 2016_TShB_severity}
TShB_sev_AUDPS <- lmer(TShB_percent_AUDPS ~ WMGT * NRTE + (1|REP), data = AUDPS_2016)
print(summary(TShB_sev_AUDPS), correlation = FALSE)

plot(TShB_sev_AUDPS)
qqnorm(resid(TShB_sev_AUDPS))
qqline(resid(TShB_sev_AUDPS))

print(lsmeans(TShB_sev_AUDPS, list(pairwise ~ NRTE)), adjust = c("tukey"))
print(lsmeans(TShB_sev_AUDPS, list(pairwise ~ WMGT)), adjust = c("tukey"))
print(lsmeans(TShB_sev_AUDPS, list(pairwise ~ WMGT | NRTE)), adjust = c("tukey"))
```

### 2016 Leaf Sheath Blight Severity Model

```{r 2016_LShB_severity}
LShB_sev_AUDPS <- lmer(LShB_percent_AUDPS ~ WMGT * NRTE + (1|REP), data = AUDPS_2016)
print(summary(LShB_sev_AUDPS), correlation = FALSE)

plot(LShB_sev_AUDPS)
qqnorm(resid(LShB_sev_AUDPS))
qqline(resid(LShB_sev_AUDPS))

print(lsmeans(LShB_sev_AUDPS, list(pairwise ~ NRTE)), adjust = c("tukey"))
print(lsmeans(LShB_sev_AUDPS, list(pairwise ~ WMGT)), adjust = c("tukey"))
print(lsmeans(LShB_sev_AUDPS, list(pairwise ~ WMGT | NRTE)), adjust = c("tukey"))
```

## Conclusions

## R Session Info

```{r session_info, echo=FALSE}
devtools::session_info()
```