### 2017-08-24 data-raw/01-Clean_2015_Raw_Data.Rmd
Add data-raw files

---
title: "Clean 2015 Raw Data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The raw data are located in the `extdata` folder of the installed
_rice.awd.pests_ and can be accessed using:

```{r, eval=FALSE}
load(system.file("extdata", "DS2015_Raw_22DAI.csv", package = "rice.awd.pests"))
load(system.file("extdata", "DS2015_Raw_35DAI.csv", package = "rice.awd.pests"))
load(system.file("extdata", "DS2015_Raw_49DAI.csv", package = "rice.awd.pests"))
load(system.file("extdata", "DS2015_Raw_62DAI.csv", package = "rice.awd.pests"))
load(system.file("extdata", "DS2015_Raw_83DAI.csv", package = "rice.awd.pests"))
```

```{r}

library(readr)
library(magrittr)

# Preprocess 2015 data for all assessment dates

files <-
  list.files("data", pattern = "^DS2015_Raw", full.names = TRUE)

reformat <- function(files) {
  x <- read_csv(
    files,
    cols(
      col_character(),
      col_character(),
      col_character(),
      col_character(),
      col_character(),
      col_character(),
      col_double(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer()
    ),
    col_names = TRUE
  )
  x[is.na(x)] <- 0

  x <- dplyr::arrange(x, REP, TRT)

  # add plot numbers
  x$PLOT <- rep(1:24, each = 18)

  # Calculate tiller sheath blight incidence -----------------------------------
  x <-
    x %>%
    dplyr::mutate(TShB_incidence = NTShB / NTIL)

  TShB_incidence <-
    x %>% dplyr::select(REP,
                        TRT,
                        WMGT,
                        NRTE,
                        SMPL,
                        HILL,
                        HGHT,
                        NTIL,
                        NTShB,
                        TShB_incidence)

  # Gather tillers--------------------------------------------------------------
  TIL <- vector(mode = "list", length = 4)

  for (i in 1:4) {
    y <-
      x %>%
      dplyr::select(
        TRT,
        REP,
        WMGT,
        SMPL,
        NRTE,
        HILL,
        paste0("TIL", i),
        paste0("GL", i),
        paste0("DL", i),
        paste0("SHB", i),
        paste0("SLA", i),
        paste0("SLB", i),
        paste0("SLC", i),
        paste0("SLD", i),
        paste0("SLE", i),
        paste0("SLF", i)
      ) %>%
      tidyr::gather(LEAF,
                    LEAF_ShB,
                    dplyr::starts_with("SL"))

    names(y)[7] <- "TIL"
    names(y)[8] <- "GL"
    names(y)[9] <- "DL"
    names(y)[10] <- "TIL_ShB"

    y$LEAF[y$LEAF == paste0("SLA", i)] <- 1
    y$LEAF[y$LEAF == paste0("SLB", i)] <- 2
    y$LEAF[y$LEAF == paste0("SLC", i)] <- 3
    y$LEAF[y$LEAF == paste0("SLD", i)] <- 4
    y$LEAF[y$LEAF == paste0("SLE", i)] <- 5
    y$LEAF[y$LEAF == paste0("SLF", i)] <- 6

    TIL[[i]] <- y
  }

  z <- dplyr::bind_rows(TIL)
  x <- dplyr::left_join(TShB_incidence, z)

  # Add dates ------------------------------------------------------------------

  switch(
    files,
    "data/DS2015_Raw_22DAI.csv" = {
      DATE <- rep(as.Date("2015-02-12", origin = "1970-01-01"),
                  times = nrow(x))
      ASMT <- rep(1, times = nrow(x))
      visit <- data.frame(DATE, ASMT, x)
    },
    "data/DS2015_Raw_35DAI.csv" = {
      DATE <- rep(as.Date("2015-02-20", origin = "1970-01-01"),
                  times = nrow(x))
      ASMT <- rep(2, times = nrow(x))
      visit <- data.frame(DATE, ASMT, x)
    },
    "data/DS2015_Raw_49DAI.csv" = {
      DATE <- rep(as.Date("2015-03-05", origin = "1970-01-01"),
                  times = nrow(x))
      ASMT <- rep(3, times = nrow(x))
      visit <- data.frame(DATE, ASMT, x)
    },
    "data/DS2015_Raw_62DAI.csv" = {
      DATE <- rep(as.Date("2015-03-19", origin = "1970-01-01"),
                  times = nrow(x))
      ASMT <- rep(4, times = nrow(x))
      visit <- data.frame(DATE, ASMT, x)
    },
    "data/DS2015_Raw_83DAI.csv" = {
      DATE <- rep(as.Date("2015-04-01", origin = "1970-01-01"),
                  times = nrow(x))
      ASMT <- rep(5, times = nrow(x))
      visit <- data.frame(DATE, ASMT, x)
    }
  )
}

# Run reformat function for all 2015 files -------------------------------------
DS2015 <- purrr::map_df(files, reformat)

# Replace "N*" with a number for NRTE ------------------------------------------
DS2015$NRTE[which(DS2015$NRTE == "N0")] <- 0
DS2015$NRTE[which(DS2015$NRTE == "N1")] <- 100
DS2015$NRTE[which(DS2015$NRTE == "N2")] <- 120

# Add treatment numbers --------------------------------------------------------
DS2015$TRT <- NA
DS2015$TRT[which(DS2015$WMGT == "FLD" &
                   DS2015$NRTE == 0)] = "FLD_N0"
DS2015$TRT[which(DS2015$WMGT == "FLD" &
                   DS2015$NRTE == 100)] = "FLD_N100"
DS2015$TRT[which(DS2015$WMGT == "FLD" &
                   DS2015$NRTE == 120)] = "FLD_N120"
DS2015$TRT[which(DS2015$WMGT == "AWD" &
                   DS2015$NRTE == 0)] = "AWD_N0"
DS2015$TRT[which(DS2015$WMGT == "AWD" &
                   DS2015$NRTE == 100)] = "AWD_N100"
DS2015$TRT[which(DS2015$WMGT == "AWD" &
                   DS2015$NRTE == 120)] = "AWD_N120"

DS2015$YEAR <- lubridate::year(DS2015$DATE)

DS2015$PLOT <- rep(1:24, each = 432)

# Arrange columns --------------------------------------------------------------
DS2015 <-
  DS2015 %>% dplyr::select(
    YEAR,
    DATE,
    ASMT,
    REP,
    PLOT,
    TRT,
    SMPL,
    HILL,
    TIL,
    LEAF,
    GL,
    DL,
    TShB_incidence,
    TIL_ShB,
    LEAF_ShB,
    WMGT,
    NRTE
  )

DS2015$LEAF_ShB <- as.numeric(DS2015$LEAF_ShB)

DS2015 <- tibble::as_tibble(dplyr::arrange(DS2015, ASMT, PLOT))

# Create new columns of days to calcluate AUDPS -------------------------------

DAYS <- as.numeric(substr(files, 17, 18))

DS2015$DAYS <- c(rep(0, 10368), rep(diff(DAYS), each = 10368))

rm(files, reformat)

# Save data for inclusion in the package ---------------------------------------

if (!dir.exists("../data")) {
  dir.create("../data", recursive = TRUE)
}

save(DS2015, file = "../data/DS2015", compress = "bzip2")

```

The raw data are located in the `extdata` folder of the installed
_rice.awd.pests_ and can be accessed using:

`load(system.file("extdata", "DS2016_Raw_1.csv", package = "rice.awd.pests"))`
`load(system.file("extdata", "DS2016_Raw_2.csv", package = "rice.awd.pests"))`
`load(system.file("extdata", "DS2016_Raw_3.csv", package = "rice.awd.pests"))`
`load(system.file("extdata", "DS2016_Raw_4.csv", package = "rice.awd.pests"))`

```{r}
# Preprocess 2016 data for all assessments
#
# 1. Cleanup date format and fill empty dates
# 2. Fill empty NTIL (for same hill, not missing values)
# 3. Fill empty NTShB (for same hill, not missing values)

# First assessment -------------------------------------------------------------
DS2016_A1 <- readr::read_csv("data/DS2016_Raw_1.csv")
# Fill missing values in NTIL and NTShB
DS2016_A1 <-
  DS2016_A1 %>% tidyr::fill(NTIL)
DS2016_A1 <-
  DS2016_A1 %>% tidyr::fill(NTShB)

# Fill dates
DS2016_A1[, 1] <- as.Date("2016-03-15", origin = "1970-01-01")
DS2016_A1$ASMT <- rep(1, nrow(DS2016_A1))
DS2016_A1$PLOT <- rep(1:16, each = 72)

# Second assessement -----------------------------------------------------------
DS2016_A2 <- readr::read_csv("data/DS2016_Raw_2.csv")
DS2016_A2 <-
  DS2016_A2 %>% tidyr::fill(NTIL)
DS2016_A2 <-
  DS2016_A2 %>% tidyr::fill(NTShB)

# Fill dates
DS2016_A2[, 1] <- as.Date("2016-03-29", origin = "1970-01-01")
DS2016_A2$ASMT <- rep(2, nrow(DS2016_A2))
DS2016_A2$PLOT <- rep(1:16, each = 72)

# Third assessement ------------------------------------------------------------
DS2016_A3 <- readr::read_csv("data/DS2016_Raw_3.csv")
DS2016_A3 <-
  DS2016_A3 %>% tidyr::fill(NTIL)
DS2016_A3 <-
  DS2016_A3 %>% tidyr::fill(NTShB)

# Fill dates
DS2016_A3[, 1] <- as.Date("2016-04-12", origin = "1970-01-01")
DS2016_A3$ASMT <- rep(3, nrow(DS2016_A3))
DS2016_A3$PLOT <- rep(1:16, each = 72)

# Fourth assessment ------------------------------------------------------------
DS2016_A4 <- readr::read_csv("data/DS2016_Raw_4.csv")
DS2016_A4 <-
  DS2016_A4 %>% tidyr::fill(NTIL)
DS2016_A4 <-
  DS2016_A4 %>% tidyr::fill(NTShB)

# Fill dates
DS2016_A4[, 1] <- as.Date("2016-04-26", origin = "1970-01-01")
DS2016_A4$ASMT <- rep(4, nrow(DS2016_A4))
DS2016_A4$PLOT <- rep(1:16, each = 72)

# Combine all assessments ------------------------------------------------------
DS2016 <- rbind(
  as.data.frame(DS2016_A1),
  as.data.frame(DS2016_A2),
  as.data.frame(DS2016_A3),
  as.data.frame(DS2016_A4)
)

# Set any NA values to 0, per Dr. Castilla on 060317 ---------------------------
# "We used to put zeroes on the data sheets, but Dr. Savary later suggested that
# we leave the cells blank because it has caused difficulties in proof reading
# the data. We also found more mistakes in data encoding when there are a lot
# of zeroes."
# Set any NA values to 0, per Dr. Castilla on 060317 ---------------------------

DS2016[is.na(DS2016)] <- 0

# Replace "Flooded" with "FLD" -------------------------------------------------
DS2016$WMGT[which(DS2016$WMGT == "Flooded")] = "FLD"

# Replace "N*" with a number for NRTE ------------------------------------------
DS2016$NRTE[which(DS2016$NRTE == "N1")] = 60
DS2016$NRTE[which(DS2016$NRTE == "N2")] = 180

# Add treatment numbers --------------------------------------------------------
DS2016$TRT <- NA
DS2016$TRT[which(DS2016$WMGT == "FLD" &
                   DS2016$NRTE == 60)] = "FLD_N60"
DS2016$TRT[which(DS2016$WMGT == "FLD" &
                   DS2016$NRTE == 180)] = "FLD_N180"
DS2016$TRT[which(DS2016$WMGT == "AWD" &
                   DS2016$NRTE == 60)] = "AWD_N60"
DS2016$TRT[which(DS2016$WMGT == "AWD" &
                   DS2016$NRTE == 180)] = "AWD_N180"

DS2016 <-
  DS2016 %>%
  tidyr::gather(
    LEAF,
    LEAF_ShB,
    dplyr::starts_with("ShBL"),
    -DATE,
    -ASMT,
    -PLOT,
    -TRT,
    -REP,
    -WMGT,
    -SMPL,
    -NRTE,
    -HILL,
    -NTIL,
    -GL,
    -DL,
    -NTShB,
    -NSHShB,
    -TIL
  )

# reclassify LEAF column to leaf numbers 1...6

DS2016$LEAF[DS2016$LEAF == "ShBL1"] <- 1
DS2016$LEAF[DS2016$LEAF == "ShBL2"] <- 2
DS2016$LEAF[DS2016$LEAF == "ShBL3"] <- 3
DS2016$LEAF[DS2016$LEAF == "ShBL4"] <- 4
DS2016$LEAF[DS2016$LEAF == "ShBL5"] <- 5
DS2016$LEAF[DS2016$LEAF == "ShBL6"] <- 6

# before classifying incidence, set any values where there are more infected
# tillers in the hill than there are tillers to NA for both columns

DS2016$NTIL[DS2016$NTIL < DS2016$NTShB] <- NA
DS2016$NTShB[is.na(DS2016$NTIL)] <- NA

# calculate tiller sheath blight incidence
DS2016 <-
  DS2016 %>%
  dplyr::group_by(DATE, ASMT, REP, PLOT, TRT, WMGT, NRTE, SMPL, HILL) %>%
  dplyr::mutate(TShB_incidence = NTShB / NTIL)

DS2016 <- dplyr::rename(DS2016, TIL_ShB = NSHShB)

DS2016$YEAR <- lubridate::year(DS2016$DATE)
DS2016$LEAF_ShB <- as.numeric(DS2016$LEAF_ShB)

DS2016 <-
  DS2016 %>% dplyr::select(
    YEAR,
    DATE,
    ASMT,
    REP,
    PLOT,
    TRT,
    SMPL,
    HILL,
    TIL,
    LEAF,
    GL,
    DL,
    TShB_incidence,
    TIL_ShB,
    LEAF_ShB,
    WMGT,
    NRTE
  )

DS2016 <- tibble::as_tibble(dplyr::arrange(DS2016, ASMT, PLOT))

# Create new columns of days to calcluate AUDPS -------------------------------

DATE_1_2016 <- DATE_2_2016 <- DS2016$DATE
DATE_1_2016[which(DATE_1_2016 == min(DATE_1_2016))] <- NA
DATE_2_2016[which(DATE_2_2016 == max(DATE_2_2016))] <- NA

DATE_1 <- na.omit(DATE_1_2016)
DATE_2 <- na.omit(DATE_2_2016)

DAYS_2016 <- lubridate::time_length(DATE_1 - DATE_2, unit = "day")
DS2016$DAYS <-
  c(rep(0, nrow(DS2016) - length(DAYS_2016)), DAYS_2016)

# cleanup workspace ------------------------------------------------------------

rm(
  DS2016_A1,
  DS2016_A2,
  DS2016_A3,
  DS2016_A4,
  DATE_1,
  DATE_1_2016,
  DATE_2,
  DATE_2_2016,
  DAYS_2016
)

# Arrange columns --------------------------------------------------------------
DS2016 <-
  DS2016 %>% dplyr::select(
    YEAR,
    DATE,
    ASMT,
    REP,
    PLOT,
    TRT,
    SMPL,
    HILL,
    TIL,
    LEAF,
    GL,
    DL,
    TShB_incidence,
    TIL_ShB,
    LEAF_ShB,
    WMGT,
    NRTE,
    DAYS
  )
```

```{r}
# Join the 2015 and 2016 Data into one Tibble ----------------------------------

RAW_data <- tibble::as_tibble(rbind(as.data.frame(DS2015),
                                    as.data.frame(DS2016)))

# convert columns to factor ----------------------------------------------------
RAW_data$YEAR <- factor(RAW_data$YEAR)
RAW_data$ASMT <- factor(RAW_data$ASMT)
RAW_data$PLOT <- factor(RAW_data$PLOT)
RAW_data$REP <- factor(RAW_data$REP)
RAW_data$TRT <- factor(RAW_data$TRT)
RAW_data$WMGT <- factor(RAW_data$WMGT)
RAW_data$NRTE <- factor(RAW_data$NRTE)
RAW_data$SMPL <- factor(RAW_data$SMPL)
RAW_data$HILL <- factor(RAW_data$HILL)
RAW_data$TIL <- factor(RAW_data$TIL)
RAW_data$LEAF <- factor(RAW_data$LEAF)

# reorder the treatments by level/year - low to high, 2015-2016 for graphs -----
RAW_data$NRTE <-
  forcats::fct_relevel(RAW_data$NRTE,
                       "0",
                       "100",
                       "120",
                       "60",
                       "180")

RAW_data$TRT <-
  forcats::fct_relevel(
    RAW_data$TRT,
    "AWD_N0",
    "AWD_N100",
    "AWD_N120",
    "AWD_N60",
    "AWD_N180",
    "FLD_N0",
    "FLD_N100",
    "FLD_N120",
    "FLD_N60",
    "FLD_N180"
  )

# add column with midpoint % tiller ShB severity -------------------------------

RAW_data <-
  dplyr::mutate(
    RAW_data,
    PERC_TIL_ShB = dplyr::case_when(
      TIL_ShB == 0 ~ 0,
      TIL_ShB == 1 ~ 1,
      TIL_ShB == 2 ~ 2.5,
      TIL_ShB == 3 ~ 10,
      TIL_ShB == 4 ~ 32.5,
      TRUE ~ 75
    )
  )

# add column with midpoint % leaf ShB severity ---------------------------------
RAW_data <-
  dplyr::mutate(
    RAW_data,
    PERC_LEAF_ShB = dplyr::case_when(
      LEAF_ShB == 0 ~ 0,
      LEAF_ShB == 1 ~ 1,
      LEAF_ShB == 2 ~ 2.5,
      LEAF_ShB == 3 ~ 10,
      LEAF_ShB == 4 ~ 32.5,
      TRUE ~ 75
    )
  )

DS2015 <- subset(RAW_data, YEAR == "2015")
DS2016 <- subset(RAW_data, YEAR == "2016")

# calculate AUDPS values -------------------------------------------------------

# 2015 AUDPS -------------------------------------------------------------------

# 2015 Tiller incidence setup --------------------------------------------------
TShB_inc_15 <-
  DS2015 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_TShB_incidence = TShB_incidence)) %>%
  dplyr::arrange(PLOT)

TShB_inc_wide <-
  reshape2::dcast(TShB_inc_15, PLOT ~ ASMT, value.var = "PLOT_TShB_incidence")

# 2015 Tiller Severity AUDPS ---------------------------------------------------
TShB_sev_15 <-
  DS2015 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_TShB_severity = TIL_ShB)) %>%
  dplyr::arrange(PLOT)

TShB_sev_wide <-
  reshape2::dcast(TShB_sev_15, PLOT ~ ASMT, value.var = "PLOT_TShB_severity")

TShB_perc_15 <-
  DS2015 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_TShB_PERCENT = PERC_TIL_ShB)) %>%
  dplyr::arrange(PLOT)

TShB_perc_wide <-
  reshape2::dcast(TShB_perc_15, PLOT ~ ASMT, value.var = "PLOT_TShB_PERCENT")

# 2015 Leaf Severity setup -----------------------------------------------------
LShB_sev_15 <-
  DS2015 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_LShB_severity = LEAF_ShB)) %>%
  dplyr::arrange(PLOT)

LShB_sev_wide <-
  reshape2::dcast(LShB_sev_15, PLOT ~ ASMT, value.var = "PLOT_LShB_severity")

LShB_perc_15 <-
  DS2015 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_LShB_PERCENT = PERC_LEAF_ShB)) %>%
  dplyr::arrange(PLOT)

LShB_perc_wide <-
  reshape2::dcast(LShB_perc_15, PLOT ~ ASMT, value.var = "PLOT_LShB_PERCENT")

# Calculate 2015 AUDPS ---------------------------------------------------------

TShB_inc_AUDPS <-
  agricolae::audps(evaluation = TShB_inc_wide[, 2:6],
                   dates = as.vector(TShB_inc_15[1:5, 6]))

TShB_sev_AUDPS <-
  agricolae::audps(evaluation = TShB_sev_wide[, 2:6],
                   dates = as.vector(TShB_sev_15[1:5, 6]))

LShB_sev_AUDPS <-
  agricolae::audps(evaluation = LShB_sev_wide[, 2:6],
                   dates = as.vector(LShB_sev_15[1:5, 6]))

TShB_percent_AUDPS <-
  audps(evaluation = TShB_perc_wide[, 2:6],
        dates = as.vector(TShB_perc_15[1:5, 6]))

LShB_percent_AUDPS <-
  audps(evaluation = LShB_perc_wide[, 2:6],
        dates = as.vector(LShB_perc_15[1:5, 6]))

AUDPS_15 <-
  tibble::as_tibble(
    cbind(
      PLOT = 1:24,
      TShB_inc_AUDPS,
      TShB_sev_AUDPS,
      TShB_percent_AUDPS,
      LShB_sev_AUDPS,
      LShB_percent_AUDPS
    )
  )

AUDPS_15$PLOT <- as.character(AUDPS_15$PLOT)
TShB_inc_15$PLOT <- as.character(TShB_inc_15$PLOT)

ShB_15 <- left_join(TShB_inc_15, AUDPS_15, by = "PLOT")

# 2015 AUDPS -------------------------------------------------------------------

# 2016 Tiller Incidence setup --------------------------------------------------
TShB_inc_16 <-
  DS2016 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_TShB_incidence = TShB_incidence)) %>%
  dplyr::arrange(PLOT)

TShB_inc_wide <-
  reshape2::dcast(TShB_inc_16, PLOT ~ ASMT, value.var = "PLOT_TShB_incidence")

TShB_sev_16 <-
  DS2016 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_TShB_severity = TIL_ShB)) %>%
  dplyr::arrange(PLOT)

TShB_sev_wide <-
  reshape2::dcast(TShB_sev_16, PLOT ~ ASMT, value.var = "PLOT_TShB_severity")

# 2016 Tiller Severity setup ---------------------------------------------------

TShB_perc_16 <-
  DS2016 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_TShB_PERCENT = PERC_TIL_ShB)) %>%
  varrange(PLOT)

TShB_perc_wide <-
  reshape2::dcast(TShB_perc_16, PLOT ~ ASMT, value.var = "PLOT_TShB_PERCENT")

# 2016 Leaf Severity setup -----------------------------------------------------

LShB_sev_16 <-
  DS2016 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_LShB_severity = LEAF_ShB)) %>%
  dplyr::arrange(PLOT)

LShB_sev_wide <-
  reshape2::dcast(LShB_sev_16, PLOT ~ ASMT, value.var = "PLOT_LShB_severity")

LShB_perc_16 <-
  DS2016 %>%
  dplyr::group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  dplyr::summarise_at(.funs = dplyr::funs(mean(., na.rm = TRUE)),
                      .vars = dplyr::vars(PLOT_LShB_PERCENT = PERC_LEAF_ShB)) %>%
  dplyr::arrange(PLOT)

LShB_perc_wide <-
  reshape2::dcast(LShB_perc_16, PLOT ~ ASMT, value.var = "PLOT_LShB_PERCENT")

# 2016 AUDPS -------------------------------------------------------------------

TShB_inc_AUDPS <-
  agricolae::audps(evaluation = TShB_inc_wide[, 2:5],
                   dates = as.vector(TShB_inc_16[1:4, 6]))

TShB_sev_AUDPS <-
  agricolae::audps(evaluation = TShB_sev_wide[, 2:5],
                   dates = as.vector(TShB_sev_16[1:4, 6]))

LShB_sev_AUDPS <-
  agricolae::audps(evaluation = LShB_sev_wide[, 2:5],
                   dates = as.vector(LShB_sev_16[1:4, 6]))

TShB_percent_AUDPS <-
  agricolae::audps(evaluation = TShB_perc_wide[, 2:5],
                   dates = as.vector(TShB_perc_16[1:4, 6]))

LShB_percent_AUDPS <-
  agricolae::audps(evaluation = LShB_perc_wide[, 2:5],
                   dates = as.vector(LShB_perc_16[1:4, 6]))

AUDPS_16 <-
  tibble::as_tibble(
    cbind(
      PLOT = 1:16,
      TShB_inc_AUDPS,
      TShB_sev_AUDPS,
      TShB_percent_AUDPS,
      LShB_sev_AUDPS,
      LShB_percent_AUDPS
    )
  )

# add plot numbers to merge with the remaing treatment data
AUDPS_16$PLOT <- as.character(AUDPS_16$PLOT)
TShB_inc_16$PLOT <- as.character(TShB_inc_16$PLOT)

ShB_16 <- left_join(TShB_inc_16, AUDPS_16, by = c("PLOT" = "PLOT"))

# Merge AUDPS data for graphing ------------------------------------------------
AUDPS <- tibble::as_data_frame(as_tibble(rbind(ShB_15, ShB_16)))
AUDPS <- tidyr::separate(data = AUDPS, col = TRT,
                         sep = "_",
                         into = c("WMGT", "NRTE"))
AUDPS <- agricolae::mutate_at(.tbl = AUDPS,
                              .funs = factor,
                              .vars = c("WMGT", "NRTE"))

# Save data for inclusion in the package ---------------------------------------

if (!dir.exists("../data")) {
  dir.create("../data", recursive = TRUE)
}

save(AUDPS, file = "../data/AWD_AUDPS", compress = "bzip2")
```