### 2018-11-04 README.Rmd
Fix badges and links (again)

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/adamhsparks/rice.awd.pests/commits/master)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-`r as.character(getRversion())`-brightgreen.svg)](https://cran.r-project.org/)
[![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/) 
[![Travis-CI Build Status](https://travis-ci.org/adamhsparks/rice_awd_pests.svg?branch=master)](https://travis-ci.org/adamhsparks/rice_awd_pests)
[![Build status](https://ci.appveyor.com/api/projects/status/mh6s2n35fgry2hjo/branch/master?svg=true)](https://ci.appveyor.com/project/adamhsparks/rice-awd-pests-9lo9u/branch/master)

## Research compendium for a report on the effects of using alternate wetting and drying irrigation techniques and nitrogen rates on sheath blight disease in rice paddies.

### Compendium DOI:

http://dx.doi.org/xxxxxxx

The files at the URL above will generate the results as found in the
publication. The files hosted at
<https://github.com/adamhsparks/rice_awd_pests> are the development
versions and may have changed since the report was published.

### Authors of this repository:

Adam H. Sparks (adam.sparks@usq.edu.au)

Nancy P. Castilla (n.castilla@irri.org)

B. Ole Sander (b.sander@irri.org)

M. Noel

### Published in:

Sparks, A,  xxxxx

### Overview of contents

This repository is our research compendium for our analysis of the effects of
alternate wetting and drying irrigation technology on rice sheath blight
disease. The compendium contains all data, code, and text associated with the
publication. The `Rmd` files in the `analysis/paper/` directory contain details
of how all the analyses reported in the paper were conducted, as well as
instructions on how to rerun the analysis to reproduce the results. The
`data/` directory in the `analysis/` directory contains all the raw data. The
`data-raw` directory contains a README.Rmd, which when knit generates the data
used in the analysis of the paper and places `.rda` files in `data/` for the
analysis.

### The supplementary files

The `data-raw` directory contains:

- a Rmd file and Markdown file detailing how the raw data was handled and
Area Under the Disease Progress Stairs (AUDPS) was calculated for analysis and
the .Rda files used in the analysis were generated

The `analysis/` directory contains:

- the manuscript as submitted (in MS Word format) and it's Rmd source file

- all the data files (in CSV file format, in the `data/` directory)

- supplementary information source files (in R markdown format) and executed
versions

- all the figures that are included in the paper (in the `figures/` directory)

### The R package

This repository is organized as an R package. There are a few R functions in
this package that were used to generate some of the diagnostic plots related to
the analysis. However, for the most part I simply used the R package
structure to help manage dependencies, to take advantage of continuous
integration for automated code testing, and so I didn't have to think too much
about how to organise the files.

To download the package source as you see it on GitHub, for offline browsing,
use this line at the shell prompt (assuming you have Git installed on your
computer):

```
git clone https://github.com/adamhsparks/rice_awd_pests.git
```

Once the download is complete, open the `rice.awd.pests.Rproj` in RStudio to
begin working with the package and compendium files.

The package has a number of dependencies on other R packages, and programs
outside of R. These are listed at the bottom of this README. Installing these
can be time-consuming and complicated, so we've used a Docker image. The Docker
image includes all the necessary software, code and data to run our analysis.
The Docker image may give a quicker entry point to the project, and is more
self-contained, so might save some fiddling with installing things.

### The Docker image

A Docker image is a lightweight GNU/Linux virtual computer that can be run as a
piece of software on Windows and OSX (and other Linux systems). To capture the
complete computational environment used for this project we have a Dockerfile
that specifies how to make the Docker image that we developed this project in.
The Docker image includes all of the software dependencies needed to run the
code in this project, as well as the R package and other compendium files. To
launch the Docker image for this project, first, 
[install Docker](https://docs.docker.com/installation/) on your computer. At the
Docker prompt, enter:

    docker run -dp 8787:8787 adamhsparks/rice-awd-pests

This will start a server instance of RStudio. Then open your web browser at
localhost:8787 or or run `docker-machine ip default` in the shell to find the
correct IP address, and log in with rstudio/rstudio.

Once logged in, use the Files pane (bottom right) open the folder for this
project, and open the `.Rproj` file for this project. Once that's open, you'll
see the `analysis/paper` directory in the Files pane where you can find the R
markdown document, and knit them to produce the results in the paper. More
information about using RStudio in Docker is avaiable at the
[Rocker](https://github.com/rocker-org)
[wiki](https://github.com/rocker-org/rocker/wiki/Using-the-RStudio-image) pages.

We developed and tested the package on this Docker container, so this is the
only platform that we're confident it works on, and so recommend to anyone
wanting to use this package to generate the vignettes, etc.

## Citation

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Reproducible Research Compendium for Analysing Effects of Water Management and Nitrogen on Rice Sheath Blight_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Licenses

Manuscript: [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

Code: [MIT](http://opensource.org/licenses/MIT) year: 
`r format(Sys.Date(), "%Y")`, copyright holder: Adam Sparks

Data: [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution
requested in reuse

## Dependencies

I used [RStudio](http://www.rstudio.com/products/rstudio/) on MacOS and Ubuntu
17.10. See the colophon section of the docx file in `analysis/paper` for a
full list of the packages that this project depends on.

## Contact

Adam Sparks, Associate Professor, Centre for Crop Health  
University of Southern Queensland, West St.  
Toowoomba, Qld 4350 Australia

(+61) 7.4631.1948   adam.sparks@usq.edu.au  
https://staffprofile.usq.edu.au/Profile/Adam-Sparks