### 2020-06-11 data-raw/README.Rmd
Clean up data-raw files

---
title: "Data Raw Files"
output:
  github_document:
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Two files in this directory detail the pre-processing of disease data, [import_and_preprocess.md](import_and_preprocess.md) and an R script, [IRRI_weather_data.R](IRRI_weather_data.R) that imports weather data for 2015 and 2016 and runs statistical checks to ensure that the dry seasons in each year are not statistically different.