### 2020-06-12 data-raw/README.Rmd
Clean up README for data-raw directory

---
title: "Data Raw Files"
output:
  github_document:
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

One file exists in this directory, [IRRI_weather_data.R](IRRI_weather_data.R).
It imports weather data for 2015 and 2016 and runs statistical checks to ensure that the dry seasons in each year are not statistically different.
They were not.